Supermartingales and Termination Analysis of Statistical PCF

### 2020-12-23 probabilisticProgrammingMartingales.tex
minor

The following theorem does not prove completeness, but is suggestive in that direction:
\begin{theorem}
For any stopping time $T$ which is almost surely finite, if $(\calF_n)_n$ is the coarsest filtration to which $T$ is adapted, then there is a supermartingale $(Y_n)_n$ adapted to $(\calF_n)_n$ and an antitone function $\epsilon$ such that $(Y_n)_n$ is an antitone ranking supermartingale with respect to $T$ and $\epsilon$.
\end{theorem}

>>>>

The following theorem does not prove completeness, but is suggestive in that direction:
\begin{theorem}
For any stopping time $T$ which is almost-surely finite, if $(\calF_n)_n$ is the coarsest filtration to which $T$ is adapted, then there is a supermartingale $(Y_n)_n$ adapted to $(\calF_n)_n$ and an antitone function $\epsilon$ such that $(Y_n)_n$ is an antitone ranking supermartingale with respect to $T$ and $\epsilon$.
\end{theorem}

### 2020-12-23 relativised-completeness.tex
minor

$y_n$ is then defined recursively by
\begin{align*}
    y_0 = 2, \quad
    y_{n+1} = (y_n - \epsilon(y_n)) \frac{\mathbb P[T > n]}{\mathbb P[T > n+1]}.
\end{align*}

@@

Although the stronger induction hypothesis was necessary for the induction to work, the only reason this is needed is to prove that $y_n > 0$ for all n, which implies that $Y_n \geq 0$. The other condition for $(Y_n)$ to be an antitone-strict supermartingale is
\begin{align*}
    & \expect{Y_{n+1} \mid \mathcal F_n} \\
    = & \expect{y_{n+1} 1_{T \geq n+1} \mid \mathcal F_n} \\
    = & y_{n+1} \expect{1_{T \geq n+1} \mid \mathcal F_n} \\ 
    = & y_{n+1} 1_{T \geq n} \frac{\mathbb P[T > n+1]}{\mathbb P[T > n]} \\
    = & (y_n - \epsilon(y_n)) 1_{T \geq n} \\
    = & Y_n - \epsilon(y_n) 1_{T \geq n}
\end{align*}
therefore $(Y_n)_n$ is an antitone strict supermartingale with respect to $(T,\epsilon)$.
\end{proof}

>>>>

The sequence $(y_n)_{n \in \omega}$ is then defined recursively by
\begin{align*}
    y_0 = 2, \quad
    y_{n+1} = (y_n - \epsilon(y_n)) \frac{\mathbb P[T > n]}{\mathbb P[T > n+1]}.
\end{align*}

@@

Although the stronger induction hypothesis was necessary for the induction to work, the only reason this is needed is to prove that $y_n > 0$ for all $n$, which implies that $Y_n \geq 0$. 
The other condition for $(Y_n)$ to be an antitone-strict supermartingale is
\begin{align*}
    & \expect{Y_{n+1} \mid \mathcal F_n} \\
    = & \expect{y_{n+1} \, 1_{\set{T \geq n+1}} \mid \mathcal F_n} \\
    = & y_{n+1} \, \expect{1_{\set{T \geq n+1}} \mid \mathcal F_n} \\ 
    = & y_{n+1} \, 1_{\set{T \geq n}} \frac{\mathbb P[T > n+1]}{\mathbb P[T > n]} \\
    = & (y_n - \epsilon(y_n)) \, 1_{\set{T \geq n}} \\
    = & Y_n - \epsilon(y_n) \, 1_{\set{T \geq n}}
\end{align*}
therefore $(Y_n)_n$ is an antitone strict supermartingale with respect to $(T,\epsilon)$.
\end{proof}
